\vspace{1.0cm} \Large \bf Integrating influenza antigenic dynamics with molecular evolution

### 2013-07-23 response.tex

The incidence data are poorly explained. In the first paragraph of the Epidemiological consequences of antigenic drift section, the authors use relative incidence. In the next paragraph, are they switching to absolute incidence? This is not clear. If they are switching to absolute incidence, I don't think they ever explain in the Methods where that data comes from. In either case, what are the justifications for using absolute versus relative incidence? Also, why do they not analyze incidence prior to 1998/1999? Finally, because each influenza season spans two calendar years, why do they not break up the data by season rather than year?

An interesting result is that estimating the serum effects improves the performance a great deal (more than including the phylogenetic data, in fact!). It suggests that there may be systematic differences between the sera. Can this be attributed to the serum itself (such as whether it is goat versus ferret), or to the particular lab or study?

Could the authors correlate the estimated serum effects back to the originating source for that serum to check? As a related point, I was surprised that inclusion of the virus effects led to only minor improvement, since prior work (http://www.ncbi.nlm.nih.gov/pubmed/19900932) suggests that some viruses are generally more resistant to HI than others. However, the virus effects are only included in a model (\#9 in Table 1) where the serum effects are already included, and it is possible to imagine that including just one of these terms somehow helps correct for both effects. Can the authors include a tenth model in Table 1 that includes the virus effects but not the serum effects? This would clarify if serum effects are actually more important than virus effects, or simply whether inclusion of either one individually is more helpful than both together.

\textbf{The inclusion of `virus effects' was meant to capture exactly this obvservation, that some viruses may be generally more resistant to HI. To further assess their importance, we followed this suggestion and included test error for a model that estimates virus effects, but leaves serum effects fixed at maximum titers. In this case, we observe similar, but not quite as substantial, improvements to test error as the model that estimates serum effects, but does not include virus effects.  From this observation, we conclude that estimating either virus or serum effects both improve the model.}

### 2013-07-23 flux.tex
Fix #31 - virus effects.

%%% errortable %%%
\begin{table}[h]
	\centering
	\caption{\textbf{Average absolute prediction error of log$_2$ HI titer for test data across models and datasets.}}
	\label{errortable}	
	\makebox[\textwidth][c]{
	\begin{tabular}{ c c c c c c c c c } 
	\hline
	\multicolumn{5}{c}{} 	&	\multicolumn{4}{c}{Test error}  \\
	Model	&	Dimen 	& Location prior	& 	Serum effects 	&	Virus effects	& 	A/H3N2	& 	A/H1N1	& 	B/Vic	& 	B/Yam	\\
	\hline	
	1		&	1D 		& Uninformed		&	Fixed 			&	None			&	1.35	&	0.94 	&	0.90	&	1.08	\\
	2		&	2D 		& Uninformed		&	Fixed 			&	None			&	0.91 	&	0.78 	&	0.82	&	0.90	\\
	3		&	3D 		& Uninformed		&	Fixed 			&	None			&	0.93 	&	0.80 	&	0.85	&	0.92	\\
	4		&	4D 		& Uninformed		&	Fixed 			&	None			&	0.98 	&	0.84 	&	0.90	&	0.97	\\	
	5		&	5D 		& Uninformed		&	Fixed 			&	None			&	1.04 	&	0.89 	&	0.98	&	1.04	\\	
	6		&	2D 		& Drift				&	Fixed 			&	None			&	0.91	&	0.75	&	0.77	&	0.83	\\
	7		&	2D 		& Diffusion/Drift	&	Fixed 			&	None			&	0.89	&	0.74	&	0.74	&	0.83	\\
	8		&	2D 		& Diffusion/Drift	&	Estimated 		&	None			&	0.77	&	0.73	&	0.66	&	0.75	\\	
	9		&	2D 		& Diffusion/Drift	&	Estimated		&	Estimated		&	0.76	&	0.71 	&	0.64	&	0.72	\\
	\hline
	\end{tabular}	
	}
\end{table}

@@

We further extend the model by estimating the strength of overall reactivity of each serum rather than fixing this at its maximum titer, and additionally, by estimating the strength of reactivity of each virus isolate.
We refer to these estimates as serum effects and virus effects, respectively.
We find that including these effects decreases test error further, ranging in improvement from 0.03 to 0.13 log$_2$ HI titers (models 8 and 9; Table \ref{errortable}).

>>>>

%%% errortable %%%
\begin{table}[h]
	\centering
	\caption{\textbf{Average absolute prediction error of log$_2$ HI titer for test data across models and datasets.}}
	\label{errortable}	
	\makebox[\textwidth][c]{
	\begin{tabular}{ c c c c c c c c c } 
	\hline
	\multicolumn{5}{c}{} 	&	\multicolumn{4}{c}{Test error}  \\
	Model	&	Dimen 	& Location prior	& 	Serum effects 	&	Virus effects	& 	A/H3N2	& 	A/H1N1	& 	B/Vic	& 	B/Yam	\\
	\hline	
	1		&	1D 		& Uninformed		&	Fixed 			&	None			&	1.35	&	0.94 	&	0.90	&	1.08	\\
	2		&	2D 		& Uninformed		&	Fixed 			&	None			&	0.91 	&	0.78 	&	0.82	&	0.90	\\
	3		&	3D 		& Uninformed		&	Fixed 			&	None			&	0.93 	&	0.80 	&	0.85	&	0.92	\\
	4		&	4D 		& Uninformed		&	Fixed 			&	None			&	0.98 	&	0.84 	&	0.90	&	0.97	\\	
	5		&	5D 		& Uninformed		&	Fixed 			&	None			&	1.04 	&	0.89 	&	0.98	&	1.04	\\	
	6		&	2D 		& Drift				&	Fixed 			&	None			&	0.91	&	0.75	&	0.77	&	0.83	\\
	7		&	2D 		& Diffusion/Drift	&	Fixed 			&	None			&	0.89	&	0.74	&	0.74	&	0.83	\\
	8		&	2D 		& Diffusion/Drift	&	Estimated 		&	None			&	0.77	&	0.73	&	0.66	&	0.75	\\	
	9		&	2D 		& Diffusion/Drift	&	Fixed 			&	Estimated		&	0.80	&	0.72	&	0.69	&	0.75	\\		
	10		&	2D 		& Diffusion/Drift	&	Estimated		&	Estimated		&	0.76	&	0.71 	&	0.64	&	0.72	\\
	\hline
	\end{tabular}	
	}
\end{table}

@@

We further extend the model by estimating the strength of overall reactivity of each serum rather than fixing this at its maximum titer, and additionally, by estimating the strength of reactivity of each virus isolate.
We refer to these estimates as serum effects and virus effects, respectively.
We find that including these effects decreases test error further, resulting in comparable improvements from estimating virus and serum effects individually and giving an improvement of between 0.03 to 0.13 log$_2$ HI titers when estimating both virus and serum effects together (models 8--10; Table \ref{errortable}).
